Propositions as Sessions, Mechanically

### 2015-04-13 paper/formalisation.tex
Tweaking the description of CPS term translation.

The translation of GV terms to CP processes is given, following
\citeauthor{Wadler:2014}, using continuation-passing style. In this style a
well-typed term in GV is parameterised by a name representing the channel
which communicates the result of the translation (a CP process). Due to the
locally nameless representation, the translation of GV terms required a
procedure to infer the type of terms in order to specify the binding type for
the corresponding CP process. An example, translating choice, is shown in
Figure~\ref{fig:trans}, unfortunately this definition is wrong because the
process P is not being opened and therefore during the proof we are unable to
reason about the typability of P opened with a sufficiently fresh variable the
reason is that P is a subprocess and therefore is not locally closed note that
I am using \citeauthor{Wadler:2014}'s style but using an LNCQ notation for
clarity. In order to construct the cut and output processes, the type of
\tm{M} and \tm{N} must be computed. So, the translation takes the environment,
in which the term is typed, as an argument to query the type of free
variables. Unfortunately, this query returns a type of \coqe!option typ!
(\coqe!option! is equivalent to \tm{Maybe} in Haskell) since it may not be
found, but in fact since only well-typed GV terms are translated it is
guaranteed to return something!  Therefore, the definition is unnecessarily
cluttered with impossible error cases.

The term translation was initially designed using a monadic approach to
account for the possible error cases in retreiving names from the
environment. The coq-ext-lib
library~\footnote{https://github.com/coq-ext-lib/coq-ext-lib} provides
definitions for monad, functor and applicative type classes and associated
instances for basic Coq types. It is similar in style to Haskell's support for
these objects. Unfortunately, this definition did not yield a suitable
function to prove the translation well-typed because the function needs to be
able to operate structurally on not only terms, but also the environment and
the processes being constructed. Consider again the translation of case. here
we require to consider nl and nr in a separate environment than m and type the
resulting case term in the combined environment we also need to open these
terms using cofinite quantification to yield a suitable induction
principle. all told the function is not readily definable. further problems do
arise as we shall see.

An alternative formulation, which avoids the need to handle error cases, is to
first define only the positive cases using an inductive type, the so-called
\textit{graph relation}, and then define a dependent function that only
considers the positive cases for which the relation is
defined~\cite{McKinna:2009}. However, several issues were encountered when
attempting to specify the translation. The graph relation requires much of the
clutter present in both the \coqe!cp_rule! and \coqe!wt_tm! definition such as
linear enviornments and cofinite quantification. the proof of cps translation
using the graph relation is a work in progress. the main difficulty is that
the representation requires opening both GV terms and CP processes in
hypotheses which would appear to duplicate a lot of the overhead of their
respective rule definitions.  again, representation is important even if the
proof has still eluded us.  while it is remains to be seen whether or not the
graph relation assists us in these matters it is nonetheless worth trying to
express a property in multiple ways

>>>>

The translation of GV terms to CP processes is given, following
\citeauthor{Wadler:2014}, using continuation-passing style. In this style a
well-typed term in GV is parameterised by a name representing the channel
which communicates the result of the translation (a CP process). Due to the
locally nameless representation, the translation of GV terms required a
procedure to infer the type of terms in order to specify the binding type for
the corresponding CP process. So, the translation takes the environment, in
which the term is typed, as an argument to query the type of free
variables. Unfortunately, this query returns a type of \coqe!option typ!
(\coqe!option! is equivalent to \tm{Maybe} in Haskell) to account for it
possibly not appearing in the environment, but in fact since only well-typed
GV terms are translated it is guaranteed to return something! Therefore, the
definition is unnecessarily cluttered with impossible error cases.

The term translation was initially designed using a monadic approach to
streamline handling of the possible error cases in retreiving names from the
environment. The coq-ext-lib
library~\footnote{https://github.com/coq-ext-lib/coq-ext-lib} provides
definitions for monad, functor and applicative type classes and associated
instances for basic Coq types. It is similar in style to Haskell's support for
these objects. Unfortunately, this definition did not yield a suitable
function to prove the translation well-typed because the function needs to be
able to operate structurally on not only terms, but also the environment and
the processes being constructed. An example, translating choice, is shown in
Figure~\ref{fig:trans}. The translation of case considers \coqe!NL! and
\coqe!NR! in a separate environment to \coqe!M! and types the resulting case
term in the combined environment. Additionally, the subterms need to be
considered as opened terms using cofinite quantification in order to yield a
suitable induction principle.

An alternative formulation, which avoids the need to handle error cases, is to
first define only the positive cases using an inductive type, the so-called
\textit{graph relation}, and then define a dependent function that only
considers the positive cases for which the relation is
defined~\cite{McKinna:2009}. However, several issues were encountered when
attempting to specify the translation. The graph relation requires much of the
clutter present in both the \coqe!cp_rule! and \coqe!wt_tm! definitions such
as linear enviornments and cofinite quantification.  Consider again
Figure~\ref{fig:trans}. unfortunately this definition is wrong because the
process P is not being opened and therefore during the proof we are unable to
reason about the typability of P opened with a sufficiently fresh variable the
reason is that P is a subprocess and therefore is not locally closed note that
I am using \citeauthor{Wadler:2014}'s style but using an LNCQ notation for
clarity. In order to construct the cut and output processes, the type of
\tm{M} and \tm{N} must be computed. the proof of cps translation using the
graph relation is a work in progress. the main difficulty is that the
representation requires opening both GV terms and CP processes in hypotheses
which would appear to duplicate a lot of the overhead of their respective rule
definitions.  again, representation is important even if the proof has still
eluded us.  while it is remains to be seen whether or not the graph relation
assists us in these matters it is nonetheless worth trying to express a
property in multiple ways
